---
title: "POLISCI 251A Homework 2 Solutions"
author: "YOUR NAME HERE"
date: "June 26, 2018"
output:
  html_document: default
  pdf_document: default
---

**Section: Write the name of your TA here**

**Collaborators: If you worked with someone in this assignment, write their names here**


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE
)
```

This assignment will analyze vote returns for California House elections and vote choice in a presidential election. The learning goals for this homework are:

- Review how to download and explore data in R
- Have a better understanding on the concept of correlation
- Create your own linear models 
- Use those models to calculate predictions
- Have a better understanding of the applications of these models 


Our specific goal in this exercise is to calculate the probability that a Democratic candidate wins a House seat in a ``swing district": one where the support for Democratic and Republican candidates is about equal and the incumbent is a Democrat.  

# 1 Preliminaries: Download data and explore variables (5 points)

In Homework 1 you loaded the data set `ca2006.csv` from your computer into R. Let's do that again. Remember, `ca2006` a slightly modified version of the 2006 House election return data from the PSCL library.

The data set contains the following variables:

- `district`: California Congressional district 
- `prop_d`: proportion of votes for the Democratic candidate for the House election
- `dem_pres_2004`: proportion of two-party presidential vote for Democratic candidate in 2004 in Congressional district
- `dem_pres_2000`: proportion of two-party presidential vote for Democratic candidate in 2000 in Congressional district
- `dem_inc`: An indicator equal to 1 if the Democrat is the incumbent
- `contested`: An indicator equal to 1 if the election is contested
- `open`:  An indicator equal to 1 if the election is contested

```{r}
### Start with a clean environment
rm(list=ls())

####  Load the 'foreign' library
library(foreign)

# set working directory to the folder where you stored your dataset
#set("YOUR PATH")
setwd("C://Users/Edgar/Dropbox/PhD/5_year/summer/ML_class/Homework/HW1/")

# Now, we can dowload the data.
ca <- read.csv("ca2006.csv") 

```


## 1.1. 

Explore the structure of the data using the function 'str()'

```{r}
# Run this
str(ca)
```

## 1.2. 

Finally, you can run a summary by using the function 'summary()'. Display the summary for 'ca'

```{r}
# Run this
summary(ca)
```


### 1.3. (5 points)

- What is the mean proportion of votes for a democratic candidate?

**Answer**
0.6220

- What percentage of districts are contested?

**Answer**
84.91 \%

- Calculate the difference between the median proportion of two-party presidential vote for Democratic candidate in Congressional district in 2000 and 2004 (You can do this by hand or with code)

**Answer**
```{r}
median(ca$dem_pres_2004)-median(ca$dem_pres_2000)
```


# 2. Correlation (25 points)

In class we studied the Pearson product-moment correlation coefficient as a measure of dependence between variables. The formula for the population is:

$$\rho_{X,Y}=Corr(X,Y)=\dfrac{Cov(X,Y)}{\sigma_X *\sigma_Y} = \dfrac{E[(X-\mu_X)*(Y -\mu_Y)]}{\sigma_X *\sigma_Y}$$
where E is the expected value operator, *cov* means covariance, $\sigma$ are the standard deviations, $\mu$ the mean expected values for each variable, and *corr* is a widely used alternative notation for the correlation coefficient.

Let's study the formula for a specific sample. We can write it as:

$$r_{X,Y}= \dfrac{\sum{(x-\bar{x})*(y -\bar{y})}}{\sqrt{(\sum(x-\bar{x})^2 *\sum(y-\bar{y})^2)}}$$

Here:
- x and y refer to vectors of variables. 
- $\bar{x}$ and $\bar{y}$ refer to the means of x and y respectively.

Using the function 'cor()' calculate the correlation matrix between the variables inside 'ca'

## 2.1. (5 points)

```{r}
## YOUR CODE HERE
cor(ca, method="pearson")
```

Note that the diagonal elements are 1. This only represent the correlation of each variable with itself. 

## 2.2. Your own correlation (20 points)

Now, let's calculate the correlation between 'dem_pres_2004' and 'dem_pres_2000' (0.9812588) by hand:

```{r}
# 1. First, create an object containing the mean for dem_pres_2000
## YOUR CODE HERE
mean_2000 <- mean(ca$dem_pres_2000)
### Now, create am object with the mean for dem_pres_2004
## YOUR CODE HERE
mean_2004 <- mean(ca$dem_pres_2004)

# 2. Second, let's calculate the numerator. In this case, the covariance.
### First, create an object that contains a vector of the diference between dem_pres_2000 and its mean
dif_2000 <- (ca$dem_pres_2000 - mean_2000)
### Second, create an object that contains a vector of the diference between dem_pres_2004 and its mean
dif_2004 <- (ca$dem_pres_2004 - mean_2004)
### Finally, sum the multiplication of these two vectors (this is the covariance)
cov_2000_2004  <- sum(dif_2000*dif_2004)


# 3.The next step is to calculate the denominator.
### Using the object containing a vector of the diference between dem_pres_2000 and its mean
### create another object that elevates this values to the power of 2
dif_2000_2 <- dif_2000^2
### Using the object containing a vector of the diference between dem_pres_2004 and its mean
### create another object that elevates this values to the power of 2
dif_2004_2 <- dif_2004^2
### Create and object called sdx that contains the multiplication of the sum of these two vectors
sdx <- sum(dif_2000_2)*sum(dif_2004_2 )


# 4. Finally, print the division of the covariance between the square root of sdx
## Corroborate that your result is exactly 0.9812588
cov_2000_2004/sqrt(sdx)
```

Note that,using the same steps, you can create your own function to calculate the correlation. 

```{r}
### YOUR CODE HERE
my_cor <- function(x,y){
         mean_x <- mean(x)
         mean_y <- mean(y)
         ## Covariance
         cov_x_y  <- sum((x-mean_x) * (y-mean_y))
         ## sdx
         sdx <- sum((x-mean_x)^2) * sum((y-mean_y)^2)
         ## Cor
         cor_x_y <- cov_x_y/sqrt(sdx)
         return(cor_x_y)
         
}
## Use your function to calculate the correlation between 'dem_pres_2004' and 'dem_pres_2000'
my_cor(ca$dem_pres_2000, ca$dem_pres_2004)
```


# 3. Plot (10 points)

Create a plot of the proportion of votes for the Democratic candidate (y axis), against the proportion of the two-party vote for the Democratic presidential candidate in 2004 (John Kerry) in the district. Be sure to clearly label the axes and provide an informative title for the plot *Hint: you can create a title with the argument 'main'*

```{r}
## YOUR CODE HERE
plot(ca$prop_d ~ ca$dem_pres_2004, pch=20,xlab = "Prop. of presidential vote for Democratic in 2004", ylab = "Prop. of votes for the Democratic candidate in House", main = "House votes versus Presidential votes")
```

# 4. Regression  (25 points)

## 4.1 (15 points)

Using the function 'lm()', regress the proportion of votes for the Democratic candidate, using the proportion of the two-party vote for the Democratic presidential candidate in 2004 in the district as predictor. Print the results using the function 'summary()'.

```{r}
## YOUR CODE HERE
lm.bi <- lm(prop_d ~ dem_pres_2004, data = ca)
## Print the results
summary(lm.bi)
```

Write the value for the intercept of this model:
**Answer**
The intercept is: -0.15390 

Write the value for the slope for dem_pres_2004 of this model:
**Answer**
The slope is: 1.38268

According to the sign of the slope, how would you describe the relationship between the two variables

**Answer**
There is a positive relationship between vote for the Democratic candidate in Congressional district in 2004 and the proportion of votes for the House election. In particular, an increase of 1\% in the vote for the Democratic candidate in Congressional district in 2004 leads to an increase of 1.38\% in the proportion of votes for the House election.

## 4.2 (10 Points)

Now, add the bivariate regression to the plot using the function 'abline()'

```{r}
## YOUR CODE HERE
plot(ca$prop_d ~ ca$dem_pres_2004, pch=20, xlab = "Prop. of presidential vote for Democratic in 2004", ylab = "Prop. of votes for the Democratic candidate in House", main = "House votes versus Presidential votes") + 
abline(lm(prop_d ~ dem_pres_2004, data = ca))
```

# 5 Cost function and predict (25 points)

## 5.1 Predict (5 points)

Using the 'predict()' function create a vector of predicted values.

```{r}
## YOUR CODE HERE
fit <- predict(lm.bi)
```


## 5.2 Cost function (5 points)

Now, let's calculate the following cost function:

$$ SSE = \sum_{i=1}^N{(Y -\hat{Y})^2}$$

This cost function is called sum of squared errors (SSE). Remember that the errors are the difference between predicted and observed values. Calculate  SSE for your model. 

```{r}
## YOUR CODE HERE
sse <- sum((ca$prop_d-fit)^2)
## Print your result
sse
```

## 5.3 Create your own function (15 points)

Using the bivariate regression and a function you have written yourself (do not use `predict()`!), report the expected vote share for the Democratic candidate if dem pres 2004 = 0.5.

```{r}
## YOUR CODE HERE
lm_pred<- function(lm_obj, indep){
  coefs<- as.numeric(lm_obj$coefficients)  
  pred <- coefs[1] + coefs[2]*indep
  return(pred)
}
### Report the expected vote  for dem_pres=0.5
lm_pred(lm.bi, .5)
```

Briefly, discuss the difference between this prediction and the ones you obtained using the 'predict()' function. *Hint: Remember our discussion about in-sample and out-of-sample predictions* 

**Answer**
The 'predict()' function produces in-sample predictions. That is, predictions obtaining using the same data used to fit the model. Since the value 0.5 is not in this sample, we consider this to be an out-of-sample prediction. 


Use your function to calculate report the expected vote share for the Democratic candidate if dem pres 2004 = 1.5.

```{r}
## YOUR answer here
lm_pred(lm.bi, 1.5)
```

Briefly, discuss the problem with this prediction and the limitations of the linear model. 

**Answer**
The problem is that a vote share of 1.5 is not a valid value. Nevertheless, the linear model maps a line for all possible values (regardless of their validity). A more accurate model would be one that restricts the possible outcomes to values between 0 and 1. 



# 7. Discussion  (10 points)

Read the following two articles:

-How I Acted Like A Pundit And Screwed Up On Donald Trump (by Nate Silver). Available here:
https://fivethirtyeight.com/features/how-i-acted-like-a-pundit-and-screwed-up-on-donald-trump/

And

- Abramowitz, A. (2016). Will time for change mean time for Trump? PS: Political Science & Politics, 49(4), 659-660. Available via Canvas. 

Write a short essay of no more than 500 words answering the following questions. Why do you think many statistical models failed to predict Trump's victory? Also consider that some models were actually able to predict his victory. Were these models more reliable or were they only obtaining the correct prediction by chance? Why do you think this? *Note: Be careful to stay within the word limit*


