---
title: "Interactive Code Lecture 3"
author: "POLISCI 251A Introduction to Machine Learning"
date: "7/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### Let's start with an clean environment
rm(list=ls())
```


## 1. Relating variables 

###1.1 

In this session we are going to introduce two new functions used to generate random numbers. 
-'set.seed()' uses a single integer argument to set a seed in your computer that will allow you to replicate a random number generator. 
-The second function is an example of random number generator from an specific distribution, in this case, the Normal distribution. 
In this chunk you are going to create two vectors of 100 numbers from a Normal distribution with mean 0 and standard deviation 1. *Hint: Again, always looking to the help window first is always advisable*

```{r}
# First, let's set the seed to 124
# YOUR CODE HERE

# Create my_x using rnorm()
# YOUR CODE HERE

# Create my_y
# YOUR CODE HERE


### NOTE: By default, rnorm() creates standard normal random variables with a mean of 0 and standard deviation of 1. However, the mean and standard deviation can be altered using the mean and sd arguments
```

Take a look to the distribution of these numbers by creating a histogram of 'my_'x using the function 'hist()'.

```{r}
# YOUR CODE HERE

```

Now, let's plot 'my_x' against 'my_y' using the function 'plot()'. In this occasion we also are going to include some additional arguments to make a nicer plot. In particular, let's specify 'pch=20' to modify the default points and  name the axis as 'Random X' and 'Random Y'. Take a look to the help window for this function first. 


```{r}
# YOUR CODE HERE

```

### 1.2 Correlation

To measure the dependence between variables there are several measures, the most common is the Pearson product-moment correlation coefficient. 

$$\rho_{X,Y}=Corr(X,Y)=\dfrac{Cov(X,Y)}{\sigma_X \sigma_Y} = \dfrac{E[(X-\mu_X)(Y -\mu_Y)]}{\sigma_X\sigma_Y}$$
where E is the expected value operator, cov means covariance, $\sigma$ are the standard deviations, $\mu$ the mean expected values for each variable, and corr is a widely used alternative notation for the correlation coefficient.

This value goes from -1 to 1. The Pearson correlation is 1 in the case of a perfect direct (increasing) linear relationship (correlation), −1 in the case of a perfect decreasing (inverse) linear relationship. If the correlation coefficient is 0 then we say that the two variables are independent. 


Using the function 'cor()' calculate the correlation between your variables. 

```{r}
# YOUR CODE HERE

```

We can see that the correlation is quite low. This is not surprising since we created these two variables independently. 

### 1.3 Creating Correlated Variables

The next step is to create two variables that are not independent. To do so, we are going to use a correlation of 0.7. Follow the code to create two correlated variables. 


```{r}
# First install the MASS package and get the library
#install.packages("MASS")
library(MASS)
 
# We will use the command mvrnorm to draw a matrix of variables
 
# Let's keep it simple, 
mu <- rep(0,2) ## Mean 0 for both variables. The'rep()' function repeats a number n times
Sigma <- matrix(.7, nrow=2, ncol=2) + diag(2)*.3 ## Correlation matrix

set.seed(124)  ### Again, set a seed before running a random generator
rawvars <- mvrnorm(n=100, mu=mu, Sigma=Sigma)

```

Take a look to the correlation between these variables:
```{r}
# YOUR CODE HERE

```

Wait! The correlation is not exactly 0.7. Why this might be happening?

**Answer:**
These variables were sampled from a multivariate normal distribution, so they only represent a larger population. We should expect that a larger sample would represent the population better and thus show a correlation closer to 0.7. 

Now, plot your two variables 

```{r}
# YOUR CODE HERE

```


**STOP!**


## 2. Setting up the time for a change model


The time-for-change model, developed by Alan Abramowitz, predicts the two-party popular vote based on the state of the economy, the incumbent’s popularity, and the time the incumbent president has been in office.

### 2.1

Download and create in object from the .csv file ``TimeChange.csv" into your computer. What class is the object you just created?

```{r}
# set your working directory first
# YOUR CODE HERE

# Create an object called 'elect' using the read.csv function
# YOUR CODE HERE

# Use the function class() to explore the class of this object
# YOUR CODE HERE

```

### 2.2

Print the dimensions of the matrix and the column names using the 'dim()' function and the 'colnames()' function.

```{r}
# YOUR CODE HERE

```

### 2.3

Provide Summary Statistics for each variable in the data set.

```{r}
# YOUR CODE HERE

```

**STOP!**


## 3. Simple Regression

### 3.1 

Our first endeavor will be to predict IncumbentVoteShare with Incumbent_Net_Approval.

Create a plot of IncumbentVoteShare against Incumbent_Net_Approval. 
```{r}
# YOUR CODE HERE
```

### 3.2

Create an object called 'mod' that contains the regression of `IncumbentVoteShare` against `Incumbent_Net_Approval`. The function to create a linear regression works as follows: lm(y~x, data='your_dataset'). Here, 'y' stands for the variable we want to predict (dependent variable) and 'x' for the predictor(s). 

```{r}
## Create the object 'mod' 
## YOUR CODE HERE

## Now, use the function summarize() to observe the results

```

### 3.3

Include a regression line of `IncumbentVoteShare` against `Incumbent_Net_Approval` on the plot. *Hint: abline might be useful*

```{r}
## YOUR CODE HERE

```

### 3.4 

Now, using the regression object you used to add the line to your plot, check the names that are provided in the `lm` object. Summarize:

1) Fitted Values
2) Residuals

And store the coefficients in a vector.

```{r}
## YOUR CODE HERE



```

**STOP!**

## 4. Prediction 

### 4.1

`predict` is an extremely useful function that you've likely worked with before. Today, we're going to write our own function using an `lm` object.

Implement a prediction function that takes as an argument an lm object and a value for net approval.

```{r}
## YOUR CODE HERE



```

### 3.2 

Report the prediction for the following values: -4, 100, -100, 50
 
```{r}
## YOUR CODE HERE

```





