---
title: "Interactive Code Lecture 7 (week 4)"
author: "POLISCI251A"
date: "July 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Intro to loops

In R there are three kinds of loops: `repeat`, `while`, and `for`. Today we are going to introduce for loops. These are used when you know how many times you want the code to repeat. The `for' accepts an iterator variable and a vector. It repeats the loop, giving the iterator each element from the vector in in turn. In general, for loops follow these steps:

1. Define a local variable as iterator and the times the loop is running with `for(i in vector)`. Note that `i` could be any other iterator. 

2. Open `{`

3. Execute your code including your local variable as iterator

4. Close `}`


Let's take a look to a simple example:

```{r}
## Print a message for squared values
for(i in 1:5){
  j <- i^2
  message("Value= ", j )
}
```

Which is exactly equivalent to: 

```{r}
## Print a message for squared values
for(x in 1:5){
  x <- x^2
  message("Value= ", x )
}
```

R for loops are particularly flexible in that they are not limited to integers, or even numbers. We can pass character vectors, logical vectors or lists:

```{r}
for(m in month.name){
   message("The month of ", m )
}
```

Now, create your own loop. First create a vector from 2010 to 2018. Using that vector, inside a loop, print the messages "The year is 2010", "The year is 2011", and so on.

```{r}
## Create your vector

## Loop

```

Some times, when we are working with loops is convenient to store our results at each iteration. To do this, we create an empty object before the loop.

For example, let's use a for loop to estimate the average of squaring the result of a roll of a die.

```{r}
### Define dice sides
nsides <- 6
#### Define trials
ntrials <- 1000
### Let's create an empty vector to store our results
## The rep function repeats a value, ntimes. In this case, 
## just missing values
trials <- rep(NA, ntrials)
### Create a for loop that in each trial executes this:
#  sample(1:nsides,1). So, basically rolls a dice
#  Hint: run just that code to see what's happening
#  In each iteration store the result in our object trials


#### Finally, using the object trials now populated with 1000 rolls
#### caculate the average of squares. 


```


## 2. Leave one cross out validation

Now let's use a for loop to perform a LOOCV. This process is easier to understand when we use quantitative predictors. Let's go back to our time for change model. Get the TimeChange data used in week 2.

```{r, eval=FALSE}
### Set your wd
#setwd('Your path')
# Create an object called 'elect' using the read.csv function
elect<- read.csv('TimeChange.csv')
```
To perform LOOCV we have to fit a model in some training data, in this case n-1, and make a prediction in the excluded observation. Then, we have to calculate the error an store in a vector. For this reason is convenient to create an empty vector first. Let's do this with a simple linear model.  

```{r, eval=FALSE}
#### Create an empty vector size nrow(elect)
errors_loocv <- rep(NA, nrow(elect))
### Loop over all rows excluding one observation in each interation
for(i in 1:nrow(elect)){
  #### Create a temporary training set by excluding row i
 
  ### Create a temprary test set with only row i
  
  ### Run a linear model predicting vote share on gdp1, gdp2, net approval and two terms
  ## Don't forget that your data is train
  
  ## Predict using this model in test
  
  ### Calculate the squared error. Error= vote share for row i minus prediction
  
  #### Store in the position i of errors_loocv
  
}

##### Calculate the mean of the test squared errors



```


